### 2017-11-14 Draft2/CopyArray.Blurb.tex
Exterminate more todos
Work

% Todo: Is 'n memory' vague?
After they are done appending to them, users often want to take dynamic arrays and convert them into raw arrays. There are multiple possible reasons for this: 1) Raw arrays hold on to exactly $n$ memory to store $n$ elements. However, dynamic and growth arrays typically use $> n$ memory so they need not grow every time $\FuncAppend$ is called. 2) A function in third-party code might only accept a raw array as an argument.

The case is even more compelling for growth arrays: 3) Since they are not entirely contiguous, they have worse data locality (even if the number of discontinuities is $O(\log n)$). 4) Raw arrays can be indexed much faster than growth arrays.

>>>>

Users often want to convert dynamic arrays to raw arrays. There are multiple reasons for this: 1) Raw arrays hold on to exactly $n$ memory to store $n$ elements. However, dynamic and growth arrays typically use more than $n$ memory, so they need not grow every time an item is appended. 2) Sometimes, functions in third-party code only accept raw arrays.

For growth arrays, the case is even more compelling: 3) Since they are not contiguous as a whole, they have worse data locality (even if the number of discontinuities is $O(\log n)$). 4) Raw arrays can be indexed much faster than growth arrays.

In this section, I will show that growth arrays can be converted to raw arrays not much slower than dynamic arrays.

### 2017-11-14 Draft2/CopyArray.Growth.tex
Make algorithms not float
Work

The growth array implementation must call $\FuncArrayCopy$ multiple times, since not all elements are contiguous.

\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncToArray$}{$\VarList$}
		\State $\LclRawArray \gets \FuncNewArray(\VarList.\FieldSize)$
		\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
			\State $\FuncArrayCopy(\LclBuffer,\ \LclRawArray,\ \LclBuffer.\FieldLength)$
		\EndFor
		\State $\FuncArrayCopy(\VarList.\FieldHead,\ \LclRawArray,\ \VarList.\FieldHeadSize)$
		\State \Return $\LclRawArray$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

>>>>

The growth array algorithm is slightly slower than the dynamic array one. It needs to loop over the tail and call $\FuncArrayCopy$ multiple times, since not all elements are contiguous. However, $n$ elements are still copied in total, so the time complexity is still $O(n)$.

\begin{algorithm}[H]
	\begin{algorithmic}
		\Function{$\FuncToArray$}{$\VarList$}
		\State $\LclRawArray \gets \FuncNewArray(\VarList.\FieldSize)$
		\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
			\State $\FuncArrayCopy(\LclBuffer,\ \LclRawArray,\ \LclBuffer.\FieldLength)$
		\EndFor
		\State $\FuncArrayCopy(\VarList.\FieldHead,\ \LclRawArray,\ \VarList.\FieldHeadSize)$
		\State \Return $\LclRawArray$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

### 2017-11-14 Draft2/Indexing.Dynamic.tex
Make algorithms not float
Add pictures, [WIP] title all the algos
Work

{\HdrNote} My algorithms do not validate any arguments; all arguments passed to them are assumed to be valid.

\begin{algorithm}
	\begin{algorithmic}[1]
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State \Return $\VarList.\FieldBuffer[\ParamIndex]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

>>>>

{\HdrNote} My algorithms assume all arguments passed to them are valid. Thus, arguments are never checked.

{\HdrNote} Although the function is called $\FuncGetItem$, the syntax other algorithms use to call it is $\VarList[\ParamIndex]$.

\begin{algorithm}[H]
	\caption{Random access \TextDynamicArray}
	\begin{algorithmic}[1]
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State \Return $\VarList.\FieldBuffer[\ParamIndex]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

### 2017-11-14 Draft2/Iterating.Dynamic.tex
Make algorithms not float
Work

The algorithm for iterating dynamic arrays is simple. It loops through all valid indices, and calls the $\FuncGetItem$ method (which is denoted $\VarList[i]$ instead of $\VarList.\FuncGetItem(i)$) for each index.

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Algorithm for ``$\TextFor \LclItem \TextIn \VarList$'', where $\VarList$ is a dynamic array}
		\For{$\LclIndex = 0,\ \LclIndex < \VarList.\FieldSize,\ \LclIndex \gets \LclIndex + 1$}
			\State $\LclItem \gets \VarList[\LclIndex]$
			\State $\ParamAction(\LclItem)$
		\EndFor
	\end{algorithmic}
\end{algorithm}

>>>>

The algorithm for iterating dynamic arrays is simple and familiar.

\begin{algorithm}[H]
	\caption{``$\TextFor \LclItem \TextIn \VarList$'' algorithm \TextDynamicArray}
	\begin{algorithmic}
		\For{$\LclIndex = 0,\ \LclIndex < \VarList.\FieldSize,\ \LclIndex \gets \LclIndex + 1$}
			\State $\LclItem \gets \VarList[\LclIndex]$
			\State $\TextDoSomethingWith \LclItem$
		\EndFor
	\end{algorithmic}
\end{algorithm}

### 2017-11-14 Draft2/Iterating.Growth.tex
Make algorithms not float
Work

The algorithm for growth arrays is not as simple, however, if the user wants optimal performance. As shown in the previous section, the algorithm for indexing growth arrays is very costly compared to its dynamic array counterpart. Thus, I wish to avoid using it in my iteration algorithm.

Fortunately, it turns out that it is possible to avoid its use:

\begin{algorithm}
	\begin{algorithmic}
		\LineComment{Algorithm for ``$\TextFor \LclItem \TextIn \VarList$'', where $\VarList$ is a growth array}
		\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
			\For{$\LclItem \TextIn \LclBuffer$}
				\State $\ParamAction(\LclItem)$
			\EndFor
		\EndFor
		\For{$\LclIndex = 0,\ \LclIndex < \VarList.\FieldHeadSize,\ \LclIndex \gets \LclIndex + 1$}
		\State $\LclItem \gets \VarList.\FieldHead[\LclIndex]$
		\State $\ParamAction(\LclItem)$
		\EndFor
	\end{algorithmic}
\end{algorithm}

>>>>

The iteration algorithm for growth arrays is not quite as simple. As shown in the previous section, growth arrays' random access algorithm is expensive; thus, it is best avoided. The following algorithm implements iteration without using the random access algorithm.

\begin{algorithm}[H]
	\caption{``$\TextFor \LclItem \TextIn \VarList$'' algorithm \TextGrowthArray}
	\begin{algorithmic}
		\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
			\For{$\LclItem \TextIn \LclBuffer$}
				\State $\TextDoSomethingWith \LclItem$
			\EndFor
		\EndFor
		\For{$\LclIndex = 0,\ \LclIndex < \VarList.\FieldHeadSize,\ \LclIndex \gets \LclIndex + 1$}
			\State $\LclItem \gets \VarList.\FieldHead[\LclIndex]$
			\State $\TextDoSomethingWith \LclItem$
		\EndFor
	\end{algorithmic}
\end{algorithm}
